\begin{thebibliography}{9}

	\bibitem{numerical} 
	Jorge Nocedal  Stephen J. Wright.
	\textit{Numerical Optimization}. Springer, (2nd Edition, 2006).

	\bibitem{PaperL-BFQS} 
	Dong C. Liu and Jorge Nocedal.
	\textit{On the limited memory BFGS method for large scale optimization}. Mathematical Programming 45 (1989), pp. 503-528.
	\\\texttt{https://people.sc.fsu.edu/\textasciitilde inavon/5420a/liu89limited.pdf}
		
	\bibitem{gradientProof} 
	Léon Bottou, Frank E. Curtis, and Jorge Nocedal.
	\textit{Optimization methods for largescale machine learning}. 
	[\textit{ Analyses of Stochastic Gradient Methods}]. Chapter 4.
	2016.
	
	\bibitem{armadillo} 		
	Conrad Sanderson and Ryan Curtin. 
	\textit{Armadillo: a template-based C++ library for linear algebra}. 
	\\\texttt{http://arma.sourceforge.net/armadillo\_joss\_2016.pdf}

	
	\bibitem{haykin} 
	Simon Haykin. 
	\textit{Neural Networks and Learning Machines}. 
	Prentice-Hall, (3rd Edition, 2008).


	\bibitem{mitchell} 
	T. M. Mitchell. 
	\textit{Machine learning}. 
	McGraw-Hill, 1997.
	
	\bibitem{goodfellow} 
	I. Goodfellow, Y. Bengio, A. Courville. 
	\textit{Deep Learning}. 
	MIT Press,  2016.
	
	\bibitem{bert03} 
	D.P. Bertsekas.
	\textit{Nonlinear Programming}. 
	Second edition. Athena Scientific, 2003.
	
	\bibitem{PaperBM} 
	A. Frangioni.
	\textit{Standard Bundle Methods: Untrusted Models and Duality}.
	\\\texttt{http://pages.di.unipi.it/frangio/abstracts.html\#NDOB18}
	
	\bibitem{PBM} 
	Marko M. M\"akel\"a, Napsu Karmitsa and Outi Wilppu.
	\textit{Proximal Bundle Method for Nonsmooth andNonconvex Multiobjective Optimization}.
	\\\texttt{http://napsu.karmitsa.fi/publications/pbm.pdf}
	
	\bibitem{NonOpt} 
	Andrzej Ruszczyński.
	\textit{Nonlinear Optimization}. Princeton University Press. 
	
	\bibitem{CBM} 
	Y. Du and A. Ruszczyński.
	\textit{Rate of Convergence of the Bundle Method}. Journal  of  Optimization Theory and Applications.
	\\\texttt{https://arxiv.org/pdf/1609.00842.pdf}


	\bibitem{monk} 
	S.B. Thrun, J. Bala, E. Boloederon and I. Bratko.
	\textit{The MONK's Problems}. 
	[\textit{A performance comparison of different learning algoritms}]. Chapter 9.
	Carnegie Mellon University CMU-CS-91-197, 1991.
	\\\texttt{https://pdfs.semanticscholar.org/94c0/418c4bd9d719e1203acfd42741ebbd343073.pdf}
	
	\bibitem{NonsmoothOPT} 
	 Bagirov, Adil M., Karmitsa, Napsu, Mäkelä, Marko M.
	\textit{Introduction to Nonsmooth optimization}. Springer link.
	\\\texttt{https://www.springer.com/gp/book/9783319081137}	

 

	
\end{thebibliography}